Evaluation of classical correlation functions from 2/3D images on CPU and GPU architectures: introducing CorrelationFunctions.jl

### 2023-09-05 paper-correlations.tex
Fix a typo
Fix formatting + run fill-region

\author[Smith]{Vasily~Postnicov}
\author[Smith,MIT]{Aleksei~Samarin}
\author[Smith]{Marina ~V. Karsanina}
\author[Aus]{Martieu ~Gravey}
\author[Smith]{Aleksey~Khlyupin}
\author[Smith]{Kirill ~M. Gerke}
\ead{kg@ifz.ru}

@@

\section{Computational methodology}

\code{CorrelationFunctions.jl} is capable to compute all aforementioned
correlation functions. Where possible we provide two implementations for each
CF. The first implementation computes a correlation function in specific
predefined directions (i.e., orientation of test vectors thrown into a sample is
fixed) \cite{jiao2014chawla,EPL1}.  There are one predefined direction in 1D
case, four directions in 2D case (two orthogonal and two diagonal) and thirteen
directions in 3D case (three orthogonal, six diagonal within each orthogonal
plane and four diagonal). All directions available are schematically represented in \cref{fig:directions}. The line segment scanning approach is slower for some
CFs but requires much less memory resources. The full map approach establishes
correlations between all points on the image is usually faster, as it leverages on Fast Fourier Transform computational speed.  Both
implementations are contained in \code{Directional} and \code{Map} modules of
the package, correspondingly.

@@

At first, we compare the three ceramic images which have somewhat different genesis depending on the mixture
ratio of powder and binder used for firing the material \cite{ceramics}. As observed from \cref{fig:real-xct}, the structure of
all three samples is visually very similar. The lineal path function supports this observation. The $S_2$ functions are
also quite similar except for Sample 2 having higher porosity that results in elevated curve (the tail for two-point probability
as expected fluctuates around porosity squared). Notably, $S_2$ and $C_2$ are almost identical - this indicates that
pore phase represents a single large cluster. Recall that for $S_2$, $C_2$ and $L_2$ functions the value at correlation length
of 0 is the porosity value. Surface functions, on the other hand, feel significant difference between samples in
how pore-solid interface is organized. This is further evidenced on pore-size and chord-length graphs, with Ceramic 2
sample with the highest porosity having larger amount of large pores. Both $P$ and $p$ show very similar behavior,
but for different correlation lengths; such behavior is attributed to quite regular pores shapes in all ceramics samples -
this is evident from much larger differences between pore-size and chord-length for SEM images considered next.

@@

The \cref{fig:real-sem} presents three 2D images of sandstones and carbonate with computed correlation functions. In this example,
unlike for quite isotropic ceramics, we evaluated CFs in the directional manner. The $S_2$ graph immediately provides us
with porosity values. It also highlights severe anisotropy in sandstone SEM images. For 2D images numerous patches of pores are
not connected to each other and, thus, $S_2$ and $C_2$ are very different. Cluster functions show how connectivity decays with
length, also we can easily observe if given pore structure percolates in a given direction - none of the images percolates
(we highlighted the clusters to make this statement easily verifiable with the naked eye). Note that $C_2$ and $L_2$ were computed
with correlations lengths equal to image dimensions explore connectivity - for other CFs this leads to noised curves due to smaller number of sampled points (for example, see correlation functions' graphs in \cite{Euras2012}).
Lineal paths are close for sandstones (with lower $L_2$ values for Sandstone 1 where large pores are occluded with clays or grain fragments),
but for carbonate with much smaller apertures of pores it is way below $L_2$ for sandstones. Also, $L_2$ suggests that computations
with correlations lengths beyond where lineal path goes to zero are not very informative. For very irregular pores on 2D SEM images,
$P$ and $p$ functions look much less similar compared to the ceramics case on \cref{fig:real-xct}. For 2D images it is also visually
easy to analyze full CF maps, as presented on \cref{fig:real-sem-map} for Sandstone 1 sample. Full correlation maps show expanded version of directional functions -
asymmetry is immediately evidenced from $S_2$, $F_{ss}$ and $F_{sv}$ maps. Moreover, the map for $C_2$ shows higher connectivity in y-direction
as compared to x-axis. The interpretation of full correlation maps is less straightforward, but anisotropy patterns in directions different from
those presented in \cref{fig:directions} can be spotted.

@@

All measurements of execution time are presented on \cref{fig:timings}. It is immediately apparent that GPU
versions (where available) are much faster than CPU versions and limited only by
amount of video memory. Functions in module \code{Directional} (which utilize
per-slice directional approach) are the slowest among other implementations, but
require less memory. Depending on CF, our video card with 24 GBs of memory could handle a cube of
~$300^3$ voxels, whilst the workstation with 512-1024 GBs of RAM can handle virtually any XCT-obtained
image volume with any correlation length.

>>>>

\author[Smith]{Vasily~Postnicov}
\author[Smith,MIT]{Aleksei~Samarin}
\author[Smith]{Marina ~V. Karsanina}
\author[Aus]{Mathieu ~Gravey}
\author[Smith]{Aleksey~Khlyupin}
\author[Smith]{Kirill ~M. Gerke}
\ead{kg@ifz.ru}

@@

\section{Computational methodology}
\code{CorrelationFunctions.jl} is capable to compute all aforementioned
correlation functions. Where possible we provide two implementations for each
CF. The first implementation computes a correlation function in specific
predefined directions (i.e., orientation of test vectors thrown into a sample is
fixed) \cite{jiao2014chawla,EPL1}.  There are one predefined direction in 1D
case, four directions in 2D case (two orthogonal and two diagonal) and thirteen
directions in 3D case (three orthogonal, six diagonal within each orthogonal
plane and four diagonal). All directions available are schematically represented
in \cref{fig:directions}. The line segment scanning approach is slower for some
CFs but requires much less memory resources. The full map approach establishes
correlations between all points on the image is usually faster, as it leverages
on Fast Fourier Transform computational speed.  Both implementations are
contained in \code{Directional} and \code{Map} modules of the package,
correspondingly.

@@

At first, we compare the three ceramic images which have somewhat different
genesis depending on the mixture ratio of powder and binder used for firing the
material \cite{ceramics}. As observed from \cref{fig:real-xct}, the structure of
all three samples is visually very similar. The lineal path function supports
this observation. The $S_2$ functions are also quite similar except for Sample 2
having higher porosity that results in elevated curve (the tail for two-point
probability as expected fluctuates around porosity squared). Notably, $S_2$ and
$C_2$ are almost identical -- this indicates that pore phase represents a single
large cluster. Recall that for $S_2$, $C_2$ and $L_2$ functions the value at
correlation length of 0 is the porosity value. Surface functions, on the other
hand, feel significant difference between samples in how pore-solid interface is
organized. This is further evidenced on pore-size and chord-length graphs, with
Ceramic 2 sample with the highest porosity having larger amount of large
pores. Both $P$ and $p$ show very similar behavior, but for different
correlation lengths; such behavior is attributed to quite regular pores shapes
in all ceramics samples -- this is evident from much larger differences between
pore-size and chord-length for SEM images considered next.

@@

The \cref{fig:real-sem} presents three 2D images of sandstones and carbonate
with computed correlation functions. In this example, unlike for quite isotropic
ceramics, we evaluated CFs in the directional manner. The $S_2$ graph
immediately provides us with porosity values. It also highlights severe
anisotropy in sandstone SEM images. For 2D images numerous patches of pores are
not connected to each other and, thus, $S_2$ and $C_2$ are very
different. Cluster functions show how connectivity decays with length, also we
can easily observe if given pore structure percolates in a given direction -
none of the images percolates (we highlighted the clusters to make this
statement easily verifiable with the naked eye). Note that $C_2$ and $L_2$ were
computed with correlations lengths equal to image dimensions explore
connectivity -- for other CFs this leads to noised curves due to smaller number
of sampled points (for example, see correlation functions' graphs in
\cite{Euras2012}).  Lineal paths are close for sandstones (with lower $L_2$
values for Sandstone 1 where large pores are occluded with clays or grain
fragments), but for carbonate with much smaller apertures of pores it is way
below $L_2$ for sandstones. Also, $L_2$ suggests that computations with
correlations lengths beyond where lineal path goes to zero are not very
informative. For very irregular pores on 2D SEM images, $P$ and $p$ functions
look much less similar compared to the ceramics case on \cref{fig:real-xct}. For
2D images it is also visually easy to analyze full CF maps, as presented on
\cref{fig:real-sem-map} for Sandstone 1 sample. Full correlation maps show
expanded version of directional functions -- asymmetry is immediately evidenced
from $S_2$, $F_{ss}$ and $F_{sv}$ maps. Moreover, the map for $C_2$ shows higher
connectivity in y-direction as compared to x-axis. The interpretation of full
correlation maps is less straightforward, but anisotropy patterns in directions
different from those presented in \cref{fig:directions} can be spotted.

@@

All measurements of execution time are presented on \cref{fig:timings}. It is
immediately apparent that GPU versions (where available) are much faster than
CPU versions and limited only by amount of video memory. Functions in module
\code{Directional} (which utilize per-slice directional approach) are the
slowest among other implementations, but require less memory. Depending on CF,
our video card with 24 GBs of memory could handle a cube of $\sim 300^3$
voxels, whilst the workstation with 512-1024 GBs of RAM can handle virtually any
XCT-obtained image volume with any correlation length.
